---
title: "Demographic Insights on Film Ratings and Preferences"
author: "Ash Munta"
date: "2024-12-04"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

```{r load_data, include=FALSE}

u_data <- read_delim("ml-100k/u.data", delim = "\t", 
                     col_names = c("user_id", "movie_id", "rating", "timestamp"))
u_user <- read_delim("ml-100k/u.user", delim = "|", 
                     col_names = c("user_id", "age", "gender", "occupation", "zip_code"))
u_item <- read_delim("ml-100k/u.item", delim = "|", 
                     col_names = c("movie_id", "title", "release_date", "video_release_date",
                                   "IMDb_URL", "unknown", "Action", "Adventure", "Animation", 
                                   "Children's", "Comedy", "Crime", "Documentary", "Drama", 
                                   "Fantasy", "Film-Noir", "Horror", "Musical", "Mystery", 
                                   "Romance", "Sci-Fi", "Thriller", "War", "Western"))

```

```{r merge_data,  include=FALSE}
merged_data <- u_data %>%
  inner_join(u_user, by = "user_id") %>%
  inner_join(u_item, by = "movie_id")

(merged_data)
```


## Research Questions

**Big Research Question:**

How do demographic factors like age, gender, and occupation influence movie ratings and genre preferences?

**Little Research Question:**

Do users from different age groups, genders, and occupations rate movies differently on average, and do they have distinct genre preferences?

## Method and Data Source

### Dataset Source:
- **The MovieLens 100k dataset**, collected by the **GroupLens Research Project** at the University of Minnesota.
- MovieLens was an **online platform** where users could rate movies and receive recommendations.
- Data Collection from **September 19, 1997**, to **April 22, 1998**.

### Data:
- 943 users provided demographic information (**age**, **gender**, **occupation**, **zipcode**).
- 100,000 ratings for 1,682 movies spanning various genres.


## Example Plot

```{r plot_data, echo = FALSE}
age_grouped_data <- merged_data %>%
  mutate(age_group = case_when(
    age < 18 ~ "Under 18",
    age >= 18 & age < 30 ~ "18-29",
    age >= 30 & age < 50 ~ "30-49",
    age >= 50 ~ "50+"
  ))

fantasy_ratings <- age_grouped_data %>%
  filter(Fantasy == 1) %>%  
  group_by(age_group) %>%  
  summarize(
    avg_rating = mean(rating, na.rm = TRUE),  
    count = n(),  
    .groups = "drop"
  )

fantasy_ratings <- fantasy_ratings %>%
  mutate(age_group = factor(age_group, levels = c("Under 18", "18-29", "30-49", "50+")))

```



```{r plot}
ggplot(fantasy_ratings, aes(x = age_group, y = avg_rating, fill = age_group)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(
    breaks = seq(1, 5, 0.5) 
  ) +
  labs(
    title = "Average Ratings for Fantasy Movies by Age Group",
    x = "Age Group",
    y = "Average Rating"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
getwd()
```


