---
title: "ChickenExample"
author: "Ash Munta"
date: "2024-11-04"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggdist)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Research Question:

How does diet impact chick growth?

## Little Research Question

Also sometimes called a "fork"

-   Descriptive: On day 21, chicks in which diet group had the highest weight?
-   As a research question: After 21 days, do the different diets leads to different weights?

## Results

```{r}
ChickWeight %>% 
  filter(Time == 21) %>%
  ggplot(aes(y= weight, x = Diet)) +
  stat_summary(geom = "bar", ##geom_bar() is the alternative, after taking the summary
               fun = mean) +
  stat_summary(geom = "errorbar",
               fun.data = mean_se,
               width = 0.2,
               color = "red")
```

## Results 2.0

```{r}
ChickWeight %>% 
  filter(Time == 21) %>%
  ggplot(aes(y = weight, x = Diet)) +
  stat_halfeye() +
  labs(title = "Results")

```

