---
title: "D2SC: Weekly Assignments"
output: html_notebook
author: "Ayush Munta"
date: "`r Sys.Date()`"  
---
# Inital Loading
```{r}
library(tidyverse)
```
# Week 1

I went to http://google.com and looked up "how to load library in R". From there, I went to images, and found an example after looking a couple photos. I checked that it was working by typing in what I saw in a chunk, and then I ran the chunk of code. 

```{r}
?tidyverse
```

Description

The 'tidyverse' is a set of packages that work in harmony because they share common data representations and 'API' design. This package is designed to make it easy to install and load multiple 'tidyverse' packages in a single step. Learn more about the 'tidyverse' at https://www.tidyverse.org.

# Assignment 2


```{r}
analogy_data <- read_csv("tidy_data/MFIndD_analogy.csv")

```
1. qualtrics_id has the unique identifier for each participant 

```{r}
num_rows <- analogy_data %>%
  summarise(total_rows = n())

num_columns <- analogy_data %>%
  summarise(total_columns = ncol(analogy_data))

num_rows
num_columns
```
2A. There are 792 rows and 6 columns. These numbers are found by running the summarise() function from dplyr in conjunction with n() or ncol(). Furthermore, these also show up while running read_csv() from the readr package or simply using dim() which is a base R package

```{r}
num_people <- analogy_data %>%
  distinct(qualtrics_id) %>%
  count()

num_people
```
2B. There are 99 people in this data set. This number is found by finding the distinct number of qualtrics_id's (which represent participants), and then using count() to count them.

```{r}
trial_counts <- analogy_data %>%
  count(qualtrics_id) %>%                    
  distinct(n)                           

trial_counts
```
2C. Everyone from the data has the same number of trials. They each have 8 in this case. This is found by using count() to return the number of trials for each participant, and then using distinct to simply by seeing how many unique value there are. In this case, since it returned one value (8) that means everyone had the same amount of trials.

# Assignment 3

```{r}
relational_summary <- analogy_data %>%
  group_by(qualtrics_id) %>%  
  summarise(relational_match_count = sum(response_type == "Rel"))  


relational_summary
```
```{r}
library(ggplot2)  
```

```{r}

ggplot(relational_summary, aes(x = relational_match_count)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Histogram of Relational Matches",
       x = "Number of Relational Matches",
       y = "Frequency") +
   scale_x_continuous(breaks = seq(0, 8, by = 1)) +
  theme_minimal()
```
It seems the extremes (0 and 8) are the most common relational matches, while 5 has none.
```{r}
wide_data <- analogy_data %>%
  select(qualtrics_id, trial_number, response_type) %>%  
  pivot_wider(names_from = trial_number, values_from = response_type)


wide_data

```

```{r, eval=FALSE}

ggplot(relational_summary) +
  geom_histogram(binwidth = 1) +
  labs(title = "Histogram of Relational Matches",
       x = "Number of Relational Matches",
       y = "Frequency")
```

I think a common error might be forgetting to have defined the x aesthetic, which is required by geom_histogram(). This will throw the error:

Error in `geom_histogram()`:
! Problem while computing stat.
Caused by error in `setup_params()`:
! `stat_bin()` requires an x or y aesthetic.

This is because the function needs to knwo what variable to use for the x or y axis.

# Assignment 4
1.
```{r}
rei_data <- read_csv("tidy_data/MFIndD_REI.csv")

head(rei_data)


str(rei_data)
```
The column type of the "response" is a chr type and "scored_response" is of the num (double) type. This is because resposne hold numerical values in the form of a character but "scored_reponse" is a double.
2.
```{r}

rei_data <- rei_data %>%
  mutate(response = str_trim(response),             
         response = str_to_lower(response),   
         response_numeric = case_when(
           response %in% c("strongly disagree", "1") ~ 1,
           response %in% c("disagree", "2")          ~ 2,
           response %in% c("neutral", "3")           ~ 3,
           response %in% c("agree", "4")             ~ 4,
           response %in% c("strongly agree", "5")    ~ 5,
           TRUE                                       ~ NA_real_
         ))

```
3.
```{r}
rei_data <- rei_data %>%
  mutate(new_scored_response = ifelse(!is.na(rev_scoring) & rev_scoring == "neg",
                                      6 - response_numeric,
                                      response_numeric))
rei_data
```
4.
```{r}
match_summary <- sum(rei_data$new_scored_response == rei_data$scored_response, 
                     na.rm = TRUE)

total_rows <- nrow(rei_data)
mismatch_summary <- total_rows - match_summary


print(paste("Number of matches:", match_summary))
print(paste("Number of mismatches:", mismatch_summary))


```
# Assignment 5
1a.
```{r}

rei_summary <- rei_data %>%
  group_by(qualtrics_id, sub_type) %>%
  summarize(score = sum(scored_response), .groups = "drop")

rei_summary

```
1b.
```{r}

na_scores <- rei_summary %>%
  filter(is.na(score))

na_scores

```
As you can see, using default NA treatment, there are NA scores. 

1c.
```{r}
rei_summary <- rei_data %>%
  group_by(qualtrics_id, sub_type) %>%
  summarize(score = sum(scored_response, na.rm = TRUE), .groups = "drop")


rei_summary

```
2.
```{r}
combined_data <- left_join(rei_summary, relational_summary, by = "qualtrics_id")

combined_data
```

```{r}

ggplot(combined_data, aes(x = relational_match_count, y = score, color = sub_type)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  
  labs(
    title = "REI Score vs. Relational Match Count",
    x = "Relational Match Count (Analogy Score)",
    y = "REI Score"
  ) +
  theme_minimal() 


```
In this scatter plot, there doesn't seem to be a strong correlation between REI score and the relational match count. Nonetheless, you can sees RR and RE tend to have higher overall scores; they also seem to grow as relational match count increases, along with EE. EA seems to have an inverse relationship.

# Assignment 6

1.

```{r}
probtask_data <- read_csv("tidy_data/MFIndD_probtask.csv")
```

2.
```{r}
distinct_conditions <- probtask_data %>%
  distinct(condition) %>% 
  pull(condition)

distinct_conditions


mean_reaction_times <- c()  
conditions <- c()           


for (i in seq_along(distinct_conditions)) {
  condition_data <- probtask_data %>%
    filter(condition == distinct_conditions[i])
  

  conditions <- c(conditions, distinct_conditions[i])
  mean_reaction_times <- c(mean_reaction_times, mean(condition_data$rt, na.rm = TRUE))
}

for (i in seq_along(conditions)) {
  cat(conditions[i], ":", mean_reaction_times[i], "\n")
}



```
There are 4 distinct values in the condition column.

3.
```{r}
# with across
probtask_summary_across <- probtask_data %>%
  group_by(condition) %>%
  summarise(
    across(c(rt, correct), 
           list(mean = ~ mean(.x, na.rm = TRUE)), 
           .names = "{col}_mean")
  )


probtask_summary_across

#  without across()
probtask_summary <- probtask_data %>%
  group_by(condition) %>%
  summarise(
    mean_reaction_time = mean(rt, na.rm = TRUE),
    accuracy = mean(correct == 1, na.rm = TRUE)
  )


probtask_summary


```

